\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-05-29 paper/ms.tex
Putting ABC talk back in

This approach is rarely usable in practice because the probability of
generating a dataset that matches the observed one is extremely low.
To work around this, we replace the proposition
$D = D_{\rm observed}$ with a logically
weaker one (i.e. one that is implied by $D = D_{\rm observed}$ but does not
imply it). This is usually done by defining a few {\em summary statistics}
$s_1(D), s_2(D), ..., s_n(D)$, which hopefully capture most of the relevant
information in the data.
Then, a discrepancy function
$\rho$ is defined, measuring how different a simulated dataset $D$ is from
the real one $D_{\rm observed}$, in terms of the summary statistics:
\begin{align}
\rho\left(\{s_1(D), ..., s_n(D)\}; \{s_1(D_{\rm observed}), ..., s_n(D_{\rm observed})\}\right)
\end{align}
The analysis then proceeds by Monte Carlo sampling the
joint posterior for $\theta$ and $D$
conditional on the logically weaker proposition
\begin{align}
\rho\left(D; D_{\rm observed}\right) < \epsilon
\end{align}
where $\epsilon$ is some small number.

@@

%The main challenges associated with ABC are:
%\begin{enumerate}[(a)]
%\item The choice of summary statistics
%\item The choice of distance function
%\item How to choose the value of $\epsilon$, or how to change it as a run
%progresses
%\item How to achieve good mixing.
%\end{enumerate}
%Challenges 1 and 2 are Bayesian in nature, i.e. they relate to the very
%definition of the posterior distribution and are well defined independent of
%the fact that we are going to use a Monte Carlo method to compute the
%results. On the other hand, challenges 3 and 4 are about the computational implementation.

@@

>>>>

This approach is rarely usable in practice because the probability of
generating a dataset that matches the observed one is extremely low.
To work around this, we replace the proposition
$D = D_{\rm observed}$ with a logically
weaker one (i.e. one that is implied by $D = D_{\rm observed}$ but does not
imply it). This is usually done by defining a few {\em summary statistics}
$s_1(D), s_2(D), ..., s_n(D)$, which hopefully capture most of the relevant
information in the data.
Then, a discrepancy function
$\rho$ is defined, measuring how different a simulated dataset $D$ is from
the real one $D_{\rm observed}$, in terms of the summary statistics:
\begin{align}
\rho\left(\{s_1(D), ..., s_n(D)\}; \{s_1(D_{\rm observed}), ..., s_n(D_{\rm observed})\}\right)
\end{align}
The analysis then proceeds by Monte Carlo sampling the
joint posterior for $\theta$ and $D$
conditional on the logically weaker proposition
\begin{align}
\rho\left(D; D_{\rm observed}\right) < \epsilon
\end{align}
where $\epsilon$ is some small number. With this proposition, the
rejection rate will still
typically be very high, but lower than with $D = D_{\rm observed}$.

\subsection{Why ABC is difficult}
The main challenges associated with ABC are:
\begin{enumerate}[(i)]
\item The choice of summary statistics $s_1(D), ..., s_n(D)$
\item The choice of distance function $\rho()$
\item How to choose the value of $\epsilon$, or how to change it as a run
progresses
\item How to make algorithms more efficient than rejection sampling.
\end{enumerate}
Challenges (i) and (ii) are Bayesian in nature, i.e. they relate to the very
definition of the posterior distribution itself.
On the other hand, challenges 3 and 4 are about the computational
implementation. Most ABC analyses are done using Sequential Monte Carlo
\citep[SMC][]{delmoral},

@@

@@

\bibitem[Del Moral et al(2012)]{delmoral}
Del Moral, P., Doucet, A. and Jasra, A., 2012. An adaptive sequential Monte Carlo method for approximate Bayesian computation. Statistics and Computing, 22(5), pp.1009-1020.
